---
title: "Transforming data set"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(stringr)
```


### Function to transform and clean dataset obtained from World Bank Open Data

The following function transforms wide-format data sets into long-format data sets obtained from [World Bank Open Data](https://data.worldbank.org/), and it also cleans the data sets to prepare them for analysis.

Last update: The function include dropping unnecessary columns and transforming "value" column (rounding to one decimal place).


```{r function}
transforming_data_wb <- function(data_new) {
  dataset_wb <- read.csv(data_new) # Importing data
  dataset_wb <- dataset_wb %>%  # Reshaping from wide-format to long-format
  pivot_longer(
    cols = colnames(dataset_wb)[5:ncol(dataset_wb)],
    names_to = "year", 
    values_to = "value"
  )
  names(dataset_wb)[c(1, 2, 3 ,4)] <- c("country", "country_code", "series", "series_code") # Formatting the variables' names
  dataset_wb$year <- str_replace_all(dataset_wb$year, ".*YR|\\.$", "") # Cleaning the "year" column
  dataset_wb$year <- as.numeric(dataset_wb$year) # Formatting the column to numeric
  dataset_wb$value <- suppressWarnings(as.numeric(dataset_wb$value))  # Formatting the column to numeric
  dataset_wb <- dataset_wb %>% drop_na(value) # Dropping NA observations
  dataset_wb <- dataset_wb[, !names(dataset_wb) %in% c("country_code", "series_code")] # Dropping unnecessary columns
  dataset_wb$value <- round(dataset_wb$value, 1) # Rounding to one decimal place
  
  return(dataset_wb) # Returning the clean dataset
}

path <- "https://raw.githubusercontent.com/GiovanadelaCruz/my472-at24-ps1-exercise1/refs/heads/main/" # Main path
transforming_data_wb(paste0(path, "world_bank_data.csv")) # Calling the function "transforming_data_wb"

```

